---
title: "Analysis"
---

# Analysis of Forest Fires in Turkey

This page presents the exploratory data analysis of forest fire statistics in Turkey between 1988 and 2024.

# Key Takeaways

- Forest fires in Turkey fluctuate significantly over time.
- Some years experienced extremely high burned forest areas.
- Burned areas and fire frequency do not always increase together.
- Negligence-related fires constitute a major proportion of forest fire causes.
- Forest fires remain an important environmental issue in Turkey.

# Load Libraries

```{r}
#| echo: true
#| code-fold: true

library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(janitor)
```

# Import and Clean Dataset

```{r}
#| echo: true
#| code-fold: true

fires <- read_excel("forest_fires.xlsx") %>%
  clean_names() %>%
  mutate(across(everything(), ~na_if(as.character(.), "-"))) %>%
  mutate(across(everything(), ~as.numeric(gsub(",", ".", .))))

glimpse(fires)
summary(fires)
```

# Number of Forest Fires by Year

```{r}
#| echo: true
#| code-fold: true

ggplot(fires, aes(x = year, y = number_of_fires)) +
  geom_line(color = "red", linewidth = 1) +
  geom_point(color = "darkred", size = 2) +
  labs(
    title = "Number of Forest Fires by Year",
    x = "Year",
    y = "Number of Fires"
  ) +
  theme_minimal()
```

The number of forest fires fluctuates considerably across years.

# Burned Forest Area by Year

```{r}
#| echo: true
#| code-fold: true

ggplot(fires, aes(x = year, y = burned_area)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_point(color = "forestgreen", size = 2) +
  labs(
    title = "Burned Forest Area by Year",
    x = "Year",
    y = "Burned Area (Hectares)"
  ) +
  theme_minimal()
```

The burned forest area changes significantly across years. The year 2021 stands out as an extreme year.

# Relationship Between Fire Numbers and Burned Area

```{r}
#| echo: true
#| code-fold: true

ggplot(fires, aes(x = number_of_fires, y = burned_area)) +
  geom_point(color = "orange", size = 3) +
  labs(
    title = "Relationship Between Fire Numbers and Burned Area",
    x = "Number of Fires",
    y = "Burned Area"
  ) +
  theme_minimal()
```

The relationship between the number of fires and burned area is not perfectly linear.

# Fire Causes Analysis

```{r}
#| echo: true
#| code-fold: true

causes <- fires %>%
  summarise(
    Intentional = sum(intentional_number, na.rm = TRUE),
    Negligence = sum(negligence_number, na.rm = TRUE),
    Natural = sum(natural_number, na.rm = TRUE),
    Unknown = sum(unknown_number, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Cause",
    values_to = "Count"
  )

ggplot(causes, aes(x = reorder(Cause, Count), y = Count, fill = Cause)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Total Forest Fires by Cause",
    x = "Cause",
    y = "Total Number of Fires"
  ) +
  theme_minimal()
```

Negligence-related fires represent one of the largest categories among forest fire causes.

# Burned Area by Fire Cause

```{r}
#| echo: true
#| code-fold: true

hectares <- fires %>%
  summarise(
    Intentional = sum(intentional_hectare, na.rm = TRUE),
    Negligence = sum(negligence_hectare, na.rm = TRUE),
    Natural = sum(natural_hecatre, na.rm = TRUE),
    Unknown = sum(unknown_hectare, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Cause",
    values_to = "Hectares"
  )

ggplot(hectares, aes(x = reorder(Cause, Hectares), y = Hectares, fill = Cause)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Burned Area by Fire Cause",
    x = "Cause",
    y = "Burned Area"
  ) +
  theme_minimal()
```

Negligence-related fires account for a substantial proportion of total burned forest areas.

# Top 10 Years with Highest Burned Area

```{r}
#| echo: true
#| code-fold: true

fires %>%
  arrange(desc(burned_area)) %>%
  slice(1:10) %>%
  ggplot(aes(x = reorder(as.factor(year), burned_area), y = burned_area)) +
  geom_col(fill = "firebrick") +
  coord_flip() +
  labs(
    title = "Top 10 Years with Highest Burned Area",
    x = "Year",
    y = "Burned Area"
  ) +
  theme_minimal()
```

2021 appears as the most destructive year in terms of burned forest area.

# Conclusion

Forest fires continue to be an important environmental issue in Turkey. The analysis shows that both fire frequency and burned area vary considerably over time. Human-related causes, especially negligence, play an important role in forest fire occurrence.


::: callout-note
Assistance was received in translating certain sections into English and in writing the code.
:::
