---
title: "Analysis"
---

# Analysis of Forest Fires in Turkey

This page presents the exploratory data analysis of forest fire statistics in Turkey between 1988 and 2024.

# Key Takeaways

- Forest fires in Turkey fluctuate significantly over time.
- Certain years experienced extremely high burned forest areas.
- Burned areas and fire frequency do not always increase together.
- Forest fires remain an important environmental issue in Turkey.

# Load Libraries

```{r}
#| echo: true
#| code-fold: true

library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(janitor)
```

# Import Dataset

```{r}
#| echo: true
#| code-fold: true

fires <- read_excel("forest_fires.xlsx") %>%
  clean_names()

names(fires)
```

# Rename Columns

```{r}
#| echo: true
#| code-fold: true

fires <- fires %>%
  rename(
    year = 1,
    burned_area = 2,
    number_of_fires = 3
  )

glimpse(fires)
summary(fires)
```

# Number of Forest Fires by Year

```{r}
#| echo: true
#| code-fold: true

ggplot(fires, aes(x = year, y = number_of_fires)) +
  geom_line(color = "red") +
  geom_point(color = "darkred") +
  labs(
    title = "Number of Forest Fires by Year",
    x = "Year",
    y = "Number of Fires"
  ) +
  theme_minimal()
```

The graph shows yearly fluctuations in the number of forest fires in Turkey.

# Burned Forest Area by Year

```{r}
#| echo: true
#| code-fold: true

ggplot(fires, aes(x = year, y = burned_area)) +
  geom_line(color = "darkgreen") +
  geom_point(color = "forestgreen") +
  labs(
    title = "Burned Forest Area by Year",
    x = "Year",
    y = "Burned Area"
  ) +
  theme_minimal()
```

The burned forest area changes significantly across years.

# Relationship Between Fire Numbers and Burned Area

```{r}
#| echo: true
#| code-fold: true

ggplot(fires, aes(x = number_of_fires, y = burned_area)) +
  geom_point(color = "orange", size = 3) +
  labs(
    title = "Relationship Between Number of Fires and Burned Area",
    x = "Number of Fires",
    y = "Burned Area"
  ) +
  theme_minimal()
```

Some years experienced fewer fires but larger burned forest areas.

# Conclusion

Forest fires continue to be an important environmental issue in Turkey. The analysis shows that both fire frequency and burned area vary considerably over time.