---
title: "Data"
---

# Data

# Data Source

The dataset used in this project was obtained from the official statistics published by the General Directorate of Forestry (OGM), the official institution responsible for forest management and forest fire monitoring in Turkey.

The official data source can be accessed from the following link:

[OGM Official Statistics](https://www.ogm.gov.tr/tr/e-kutuphane/resmi-istatistikler)

# About the Dataset

This dataset contains yearly forest fire statistics in Turkey between 1988 and 2024.

The dataset includes the following variables:

- `Year`: Year of observation
- `Burned_area`: Total burned forest area in hectares
- `Number_of_fires`: Total number of forest fires
- `Intentional_number`: Number of intentional fires
- `Intentional_hectare`: Burned area caused by intentional fires
- `Negligence_number`: Number of negligence-caused fires
- `Negligence_hectare`: Burned area caused by negligence
- `Natural_number`: Number of natural fires
- `Natural_hecatre`: Burned area caused by natural fires
- `Unknown_number`: Number of fires with unknown causes
- `Unknown_hectare`: Burned area caused by unknown reasons

The data was selected because forest fires are an important environmental and socio-economic issue in Turkey. Analyzing these statistics helps us understand long-term trends, fire severity, and the role of different fire causes.

# Importing the Dataset

```{r}
#| echo: true
#| code-fold: true

library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)

fires <- read_excel("forest_fires.xlsx") %>%
  clean_names()

head(fires)
```

# Data Preparation

The original dataset contains some missing values represented with `-`, especially for fire cause variables before 1997. These values were converted into `NA`.

Additionally, column names were standardized and numeric values were converted into a suitable format for analysis.

```{r}
#| echo: true
#| code-fold: true

fires_clean <- fires %>%
  mutate(across(everything(), ~na_if(as.character(.), "-"))) %>%
  mutate(across(everything(), ~as.numeric(gsub(",", ".", .))))

glimpse(fires_clean)

summary(fires_clean)
```

# Final Dataset

The cleaned dataset was saved in `.RData` format for reproducibility.

```{r}
#| echo: true
#| code-fold: true

save(fires_clean, file = "final_forest_fires.RData")
```

The final cleaned dataset can be downloaded here:

[Download Final Dataset](final_forest_fires.RData)